\begin{solution}{Question 7}\label{ques:7}
    \begin{question}
     Give an example of a language that is not a CFL but nevertheless acts like a CFL in the pumping lemma for CFL (Recall we saw such an example in class while studying pumping lemma for regular languages). 
    \end{question}
    \tcblower{}
    \begin{proof}[Solution]
        While studying pumping lemma for regular languages, we came across the example: 
        \begin{equation}
            \begin{split}
                L_1 &= \{ ca^n b^n | n \geq 1\}\\
                L_2 &= \{ c^kw | k\neq 1, \text{ w starts with a or b}\}\\
                L &= L_1 \cup L_2
            \end{split}
        \end{equation}
        From this example, we gathered that despite $L_1$ and subsequently $L$ being non regular, L is still pumpable as the non regular part can be pumped to the regular part.\par
        Similar to this, an example of a language which is not a CFL but can still be pumped using the pumping lemma for CFLs is:
        \begin{equation}
            \begin{split}
                L_1 &= \{ d a^n b^n c^n | n \geq 1\}\\
                L_2 &= \{ d^kw | k\neq 1, \text{ w starts with a or b or c}\}\\
                L &= L_1 \cup L_2
            \end{split}
        \end{equation}
        
        $L$ is not a CFL but it can be pumped. The proof is as follows:
        
        \begin{proof}[1. L is not a CFL]
            $L_1$ is clearly not a CFL (we have proved this in the class) whereas $L_2$ is a CFL. Also, $L_1 \cap L_2 = \phi$. Therefore, it is not possible that the union of these two languages will yield a CFL. No automaton will be be able to recognize languages that begin with $d$ followed by $a, b, c$.
        \end{proof}
        
        \begin{proof}[2. L can be pumped]
        Let the pumping length be $p$ and consider a string $l \in L$ of size $\geq p$. As stated previously, $L_1$ and $L_2$ have no intersection. Therefore, case 1: \textit{$l\in L_1$}, we can pump the terminal letter $d$ ($u = \phi, v = d, w = \phi, x = \phi, y = a^nb^nc^n$) and it will fall in $L_2$ satisfying the pumping lemma; case 2: \textit{$l \in L_2$}, again, the first letter $d$ can be pumped and the resultant string will again lie in $L_2$, making it pumpable.
        
        \end{proof}
        
        Hence $L$ is an example of a language which is pumpable but not a CFL.
        
        \end{proof}
\end{solution}
