\begin{solution}{Question 1}\label{ques:1}
    \begin{question}
        Show that every infinite Turing-recognizable language has an infinite decidable subset.
    \end{question}
    \tcblower{}
    \begin{proof}
      Let the Turing-recognizable language be $L$. Consider an enumerator for this language $E$. Consider the following algorithm:
      \begin{algorithm}[H]
        \caption{Return the first word of length $n$ enumerated such that all words enumerated earlier are of shorter length else $\epsilon$}
        \label{alg:1.1}
        \begin{algorithmic}[1]
          \Procedure{EnumerateFirst}{$E, n$}
            \State{$w \gets \epsilon$}
            \State{$curr_w \gets \epsilon$}
            \While{$|curr_w| < n$}
              \State{$curr_w \gets next(E)$}
              \If{$|curr_w| > |w|$}
                \State{$w \gets curr_w$}
              \EndIf{}
            \EndWhile{}
            \If{$|w| > n$}
              \State{$w \gets \epsilon$}
            \EndIf{}
            \State{\Return{$w$}}
          \EndProcedure{}
        \end{algorithmic}
      \end{algorithm}
      We now define the following Turing machine $T_{decidable}$:
      \begin{algorithm}[H]
        \caption{Turing machine of a decidable subset}
        \label{alg:1.2}
        \begin{algorithmic}[1]
          \Procedure{DecidableSubset}{$L, w$}
            \State{$E \gets enumerator(L)$}
            \State{$w' \gets EnumerateFirst(E, |w|)$}
            \If{$w' = w$}
              \State{\Return{accept}}
            \Else{}
              \State{\Return{reject}}
            \EndIf{}
          \EndProcedure{}
        \end{algorithmic}
      \end{algorithm}
      We claim that $T_{decidable}$ is a Turing machine and decidable. For this, we first prove that Algorithm~\ref{alg:1.1} is correct and terminates. For correctness, notice that the while loop exits only if $|curr_w| \geq n$ and thus the length of the string in $w$ after the termination of the while loop is at least $n$. Therefore, the final returned value is the first string enumerated with length equal to $n$ (such that all strings before are of smaller length) or $\epsilon$.\par
      For termination, notice that there are only a finite possibilities before $E$ enumerates a string with length $\geq n$. Therefore, the while loop terminates after a finite number of iterations.\par
      We have shown that Algorithm~\ref{alg:1.1} is decidable since it halts. Therefore, the string $w'$ in ALgorithm~\ref{alg:1.2} is generated in finite number of steps. The comparison of $w'$ and $w$ also completes in finite steps. Therefore, Algorithm~\ref{alg:1.2} accepts or rejects $w$ in finite number of steps. Therefore, $T_{decidable}$ is a decidable Turing machine.\par
      We also need to show that $T_{decidable}$ recognizes an infinite language. This is obvious since $L$ is infinite and therefore the following set forms an infinite sequence:
      \begin{equation}
        \{i | \forall i \in \mathbb{N}: EnumerateFirst(E, i) \neq \epsilon\}
      \end{equation}
      This set is a bijection to the words recognised by $T_{decidable}$. Therefore, $L(T_{decidable})$ is also an infinite set.\par
      Thus, we have shown that for every infinite Turing-recognizable language, we have an infinite decidable subset.
    \end{proof}
\end{solution}
