\begin{solution}{Question 3}\label{ques:3}
    \begin{question}
        Let $C$ be a language. Prove that C is Turing-recognizable iff a decidable language $D$ exists such that
        $$C = \{ x | \exists y (\langle x,y \rangle \in D) \}$$
    \end{question}
    \tcblower{}
    \begin{proof}
      To prove the iff statement, we consider the following parts one by one:
      \begin{proof}[Direction 1:]
        Given a decidable language $D$, the Language $C = \{ x | \exists y (\langle x,y \rangle \in D) \}$ formed is Turing recognizable.\par
        \textbf{Proof:}
        For a language to be Turing recognizable, if a string $x \in \Sigma^*$ the TM halts and accepts it and if $x \not\in \Sigma^*$, TM either rejects it or keeps running.\\
        In our case, Language $D$ is Turing decidable, which means given an input $\langle x,y \rangle$ it either accepts or rejects the string and halts for every input.\\
        Given any string $x$, $C$ accepts it if there exists a string $y$ such that $\langle x,y \rangle \in D$. Hence our Turing Machine for $C$ keeps searching for a string $y$ given an input $x$ so that $\langle x,y \rangle \in D$. Since $y$ is of finite length it can always be found in finite amount of time, if such $y$ is found, $x$ is an accepted string and the TM halts. If no such $y$ is found, TM keeps looking for it indefinitely.\par
        This implies that language $C$ is Turing recognizable
      \end{proof}
      
      \begin{proof}[Direction 2:]
        Given language $C$ such that $C$ is Turing recognizable, then there exists a Turing-decidable language $D$ such that $C = \{ x | \exists y (\langle x,y \rangle \in D) \}$.\par
        \textbf{Proof:}
        We define the following algorithm for the language $D$:
        \begin{algorithm}[H]
          \caption{Algorithm for a language $D$}
          \label{alg:3.1}
          \begin{algorithmic}[1]
            \Procedure{D}{$C, x, y$}
              \State{$status \gets run_n(C, x, |y|)$}  \Comment{$run_n(M, w, n)$ simulates $M$ with input $w$ for \textit{atmost} $n$ steps}
              \If{$status = accepted$}
                \State{\Return{accept}}
              \Else{}
                \State{\Return{reject}}
              \EndIf{}
            \EndProcedure{}
          \end{algorithmic}
        \end{algorithm}
        Since Algorithm~\ref{alg:3.1} simulates $C$ on $x$ for \textit{atmost} $|y|$ iterations, the algorithm always terminates in finite number of steps. Also, it uniquely accepts or rejects $\langle x, y\rangle$. Therefore, the language $D$ is decidable. Also, it is easy to see that the language satisfies the property that $C$ accepts $x$ iff $\exists y: \langle x, y\rangle \in D$. Therefore, we have shown the existence of such a language.
      \end{proof}
      
      On combining both proven statements in Direction 1 and Direction 2, we can say that given language $C = \{ x | \exists y (\langle x,y \rangle \in D) \}$ is Turing recognizable if language $D$ is Turing Decidable.
      \begin{center}
          Hence Proved
      \end{center}
    \end{proof}
\end{solution}
