\begin{solution}{Question 5}\label{ques:5}
    \begin{question}
        Consider the problem of determining whether a Turing machine $M$ on an input $w$ ever attempts to move its head left when its head is on the left-most tape cell. Formulate this problem as a language and show that it is undecidable. 
    \end{question}
    \tcblower{}
    \begin{proof}[Solution]
        Consider the language $L_{TM}$ defined as
        \begin{equation}
            \begin{split}
                L_{TM}  &= \{\langle M, w\rangle | \\
                        &M \text{ attempts to move its head left when its head is on the left most cell on input } w\}
            \end{split}
        \end{equation}
        
        Let's assume that the language $L_{TM}$ is decidable and Turing machine $R$ decides it. $R$ takes $\langle M, w\rangle$ as input and accept if $M$'s tape head attempts to move left when it is on the left most tape cell, else it rejects. Now, consider the following Turing machine $S$ for language $A_{TM}$. It works as:
        
        \bigskip
        
        $S$ = "On input $\langle M, w\rangle$, where $M$ is a Turing machine and $w$ is a string:
        \begin{enumerate}
            \item Construct Turing machine $M_1$ using $M$ and $w$ according to definition given below.
            \item Simulate $R$ on input $\langle M_1, w\rangle$
            \item If $R$ accepts then accept else reject.
        \end{enumerate}
        
        $M_1$ = "On input $w$:
        \begin{enumerate}
            \item Make left most symbol on tape as $"\#"$ and move input string right by one cell on tape.
            \item Simulate $M$ on $w$ placing head on second symbol(i.e start of input string) as initial configuration. While simulating $M$ it doesn't allows head to move left of first symbol i.e first cell behaves as start of tape for $M$.
            \item If $M$ accepts then try moving head to left of leftmost symbol and accept; else reject.
        \end{enumerate}
        
        Here, $M_1$ attempts to move its head left of left most cell on tape iff $M$ accepts $w$. This means $R$ only accepts $M_1$ iff $M$ accepts $w$ else it rejects it. Thus we can say that $S$ decides $A_{TM}$. But this is contradiction to the fact that $A_{TM}$ is undecidable language. Our assumption that $L_{TM}$ is decidable is false. \\
        
        Hence, $L_{TM}$ is undecidable.
        
    \end{proof}
\end{solution}
