\begin{solution}{Question 5}\label{ques:5}
	\begin{question}
		For any string $w = w_1w_2\ldots w_n$ the reverse of $w$ written $w^R$ is the string $w_n\ldots w_2w_1$. For any language $A$, let $A^R = \{w^R | w\in A\}$. Show that if $A$ is regular, then so is $A^R$. In other words, regular languages are closed under the reverse operation.
	\end{question}
	\tcblower{}
	\begin{proof}
		Consider the DFA for $D_A = (Q, \Sigma, \delta, q_0, F)$. Construct the following NFA $N_A = (Q', \Sigma', \delta', q', F')$:
		\begin{equation}
			\begin{split}
				Q' &= Q \cup q'\\
				\Sigma' = \Sigma\\
				\delta'(q_i, a) &= q_j \iff \delta(q_j, a) = q_i\\
				\delta'(q', \epsilon) &= F\\
				F' &= \{q_0\}
			\end{split}
		\end{equation}
		\begin{claim}
			$N_A$ recognises the language $A^R$, i.e., every word in $A^R$ is accepted by $N_A$ and every word not in $A^R$ is rejected by $N_A$
		\end{claim}
		\begin{proof}
			Consider any word $w$ in $A$, it is recognised by $D_A$ (from construction). Consider the sequence of states visited during the run of acceptance of $w$. Let it be $S = \{q_0=q_{w_1}, q_{w_2}, \ldots, q_{w_n}\}$. Now consider the following run in $N_A$ of $w^R$:
			\begin{equation}
				\{q', q_{w_n}, \ldots, q_{w_2}, q_0=q_{w_1}\}
			\end{equation}
			The above is a valid run since each transition is a valid one from the construction of $N_A$ (the first one is an $\epsilon$-transition and the remaining are fixed transitions). This proves that every word in $A$ is recognised by $N_A$.\par
			To prove the converse, assume by contradiction that a word $w^R = w_n\ldots w_2w_1 \not\in A$ is accepted by $N_A$. Let the sequence of states be $S' = \{q', q_{w_n}, \ldots, q_{w_2}, q_{w_1}=q_0\}$. Now, from the construction of $N_A$, the following sequences of states should be accepted by $D_A$:
			\begin{equation}
				\{q_{w_1}=q_0, q_{w_2}, \ldots, q_{w_n}\}
			\end{equation}
			This corresponds to accepting the string $w$. However, $D_A$ accepts only strings in $A$. This is a contradiction to our assumption. Therefore, $N_A$ rejects every string not in $A$.\par
			Therefore, $A^R$ is recognised by $N_A$ which is an NFA. Therefore, $A^R$ is a regular language as well. Thus, regular languages are closed under the reverse operation. Hence, proved.
		\end{proof}
	\end{proof}
\end{solution}
